---
title: "P8105_hw3_yx2510"
author: "Yi Xiao"
date: "06/10/2018"
output: html_document
---
```{r}
library(tidyverse)
library(magrittr)
library(hexbin)
```
# problem one
load the data
```{r}
brfss_data = p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  rename(state = locationabbr, county_state = locationdesc) %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(response = as.factor(response))
  
brfss_data
```
### questions
#### 1. In 2002, which states were observed at 7 locations?
```{r}
brfss_data %>% 
  group_by(state) %>%
  summarize(location_num = n_distinct(geo_location)) %>%
  filter(location_num == 7)

```
There are seven states satisfy the standard, which are CT, MT, NH, NM, OR, TN and UT.

#### 2. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
brfss_data %>% 
         group_by(state, year) %>% 
         summarize(n_observ = n()) %>% 
         ggplot(aes(x = year, y = n_observ)) + 
         geom_line(aes(color = state)) + 
         labs(
                 x = "Year", 
                 y = "Number of observations") + 
          theme(
                  legend.position = "bottom"
               ) + 
          guides(color = guide_legend(ncol = 20))
```

#### 3. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
filter(brfss_data, 
       state == "NY",
       year %in% c(2002, 2006, 2010),
       response == "Excellent") %>%
  group_by(year) %>% 
  summarize(excel_sd = sd(data_value),
            excel_mean = mean(data_value)) %>% 
  knitr::kable(digits = 2)
```

#### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss_data %>%
        group_by(year, state, response) %>%
        summarise(response_mean = mean(data_value)) %>%
        ggplot(aes(x = year, y = response_mean)) +
        geom_line(aes(color = state)) +
        facet_grid(. ~ response) +
        theme(legend.position = "bottom")  + 
          guides(color = guide_legend(ncol = 20))
```

# problem 2
load and clean the data
```{r}
instacrat_data = p8105.datasets::instacart %>%
        janitor::clean_names() 
head(instacrat_data)
```
This dataset contains `r nrow(instacrat_data)` observations and `r ncol(instacrat_data)` columns. Some key variables include product_id, product_name, aisle.

## questions
#### 1.How many aisles are there, and which aisles are the most items ordered from?
```{r}
n_distinct(instacrat_data$aisle, na.rm = TRUE)
```
There are 134 distinct aisles.
```{r}
instacrat_data %>%
        group_by(aisle) %>%
        summarise(aisle_num = n()) %>%
        filter(aisle_num == max())
```
Fresh vegetables is the aisle with the most item bought from.

#### 2. Make a plot that shows the number of items ordered in each aisle. Order aisles sensibly, and organize your plot so others can read it.
Since 134 aisles are contained in the dataset, we first divide them into seveal groups for better display. 
```{r}
    # group the dataset by aisle and rearrange in descendant order

plot_aisle_num = function(left, right) {  
        instacrat_data %>%
        group_by(aisle) %>%
        summarise(aisle_num = n()) %>%
        arrange(desc(aisle_num)) %>%
        filter(aisle_num >= left & aisle_num < right) %>%
        ggplot(aes(x = aisle, y = aisle_num)) +
        geom_bar(stat = "identity") +
        geom_text(aes(label = aisle_num), size = 2.5, position = position_dodge(width = 0.9), angle = 0, vjust = - 0.5) +
        theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
        labs(
                 x = "Aisle", 
                 y = "Number of Items") 
        
}

plot_aisle_num (0, 1000)
plot_aisle_num (1000, 2000)
plot_aisle_num (2000, 5000)
plot_aisle_num (5000, 10000)
plot_aisle_num (10000, 15000)
plot_aisle_num (15000, 151000)

```


#### 3.Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”
```{r}
instacrat_data %>%
        filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
        group_by(aisle, product_name) %>%
        summarise(pro_num = n()) %>%
        filter(pro_num == max(pro_num)) %>%
        knitr::kable()
  
```


#### 4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).
```{r}
instacrat_data %>%
        filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
        group_by(product_name, order_dow) %>% 
        summarize(mean_hour = round(mean(order_hour_of_day))) %>%
        spread(key = order_dow, value = mean_hour) %>%
        knitr::kable()
                    
```

# problem 3 
load the data first 
```{r}
noaa_data = p8105.datasets::ny_noaa
str(noaa_data)
head(noaa_data)
```
The dataset has `r nrow(noaa_data)` observations and `r ncol(noaa_data)` variables including information on date, precipitation, snow and snow depth, maximum and minimum temperature. All of the variables have missing values except id and date. ????issue

## questions
#### 1. Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
```{r}
noaa_tidy_data =
        separate(noaa_data, date, into = c("year", "month", "day"), sep = "-") %>%
        mutate(prcp = prcp/10,
               snow = snow/10, 
               tmin = as.integer(tmin)/10,
               tmax = as.integer(tmax)/10,
               year = as.integer(year),
               month = as.integer(month),
               day = as.integer(day)
        ) 
noaa_tidy_data

noaa_tidy_data %>%
        group_by(snow) %>%
        summarise(snow_frequency = n()) %>%
        arrange(desc(snow_frequency))
```


#### 2.Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
noaa_tidy_data %>%
        filter(month == 1 | month == 7) %>%
        group_by(id, year, month) %>%
        summarise(tmax_mean = round(mean(tmax, na.rm = TRUE), 2), 
                  tmin_mean = round(mean(tmin, na.rm = TRUE), 2)) %>%
        gather(key = mean, value = mean_temperature, tmax_mean:tmin_mean) %>%
        ggplot(aes(x = as.factor(year), y = mean_temperature, fill = mean)) +
        geom_boxplot() +
        facet_grid(~ month) +
        theme(legend.position = "bottom") +
        theme(axis.text.x = element_text(size = 7, angle = 45)) +
        labs( x = "year",
              y = "average of maximum and minimum teperature in Januray and July")
```


#### 3.Make a two-panel plot showing: 
(i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option).
```{r}
noaa_tidy_data %>%
        ggplot(aes(x = tmin, y = tmax)) +
        geom_hex()
        
```
(ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
```{r}
noaa_tidy_data %>%
        filter(snow > 0 & snow < 100) %>%
        ggplot(aes(x = as.factor(year), y = snow)) +
        geom_boxplot() +
        labs(
                x = "Year",
                y = "Snowfall between 0 and 100"
        ) +
        theme(axis.text.x = element_text(size = 7, angle = 45))
```

