---
title: "P8105_hw3_yx2510"
author: "Yi Xiao"
date: "06/10/2018"
output: html_document
---
```{r}
library(tidyverse)
library(magrittr)
```
# problem one
load the data
```{r}
brfss_data = p8105.datasets::brfss_smart2010 %>% 
  janitor::clean_names() %>%
  rename(state = locationabbr, county_state = locationdesc) %>%
  filter(topic == "Overall Health") %>%
  filter(response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")) %>%
  mutate(response = as.factor(response))
  
brfss_data
```
### questions
#### 1. In 2002, which states were observed at 7 locations?
```{r}
brfss_data %>% 
  group_by(state) %>%
  summarize(location_num = n_distinct(geo_location)) %>%
  filter(location_num == 7)

```
There are seven states satisfy the standard, which are CT, MT, NH, NM, OR, TN and UT.

#### 2. Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
brfss_data %>% 
         group_by(state, year) %>% 
         summarize(n_observ = n()) %>% 
         ggplot(aes(x = year, y = n_observ)) + 
         geom_line(aes(color = state)) + 
         labs(
                 x = "Year", 
                 y = "Number of observations") + 
          theme(
                  legend.position = "bottom"
               ) + 
          guides(color = guide_legend(ncol = 20))
```

#### 3. Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State.
```{r}
filter(brfss_data, 
       state == "NY",
       year %in% c(2002, 2006, 2010),
       response == "Excellent") %>%
  group_by(year) %>% 
  summarize(excel_sd = sd(data_value),
            excel_mean = mean(data_value)) %>% 
  knitr::kable(digits = 2)
```

#### For each year and state, compute the average proportion in each response category (taking the average across locations in a state). Make a five-panel plot that shows, for each response category separately, the distribution of these state-level averages over time.

```{r}
brfss_data %>%
        group_by(year, state, response) %>%
        summarise(response_mean = mean(data_value)) %>%
        ggplot(aes(x = year, y = response_mean)) +
        geom_line(aes(color = state)) +
        facet_grid(. ~ response) +
        theme(legend.position = "bottom")  + 
          guides(color = guide_legend(ncol = 20))
```

# problem 2
load and clean the data
```{r}
brfss_data = p8105.datasets::instacart %>%
        janitor::clean_names() 
```

